---
title: "Project_3"
author: "Jiayu Tan 300102805"
date: "`r Sys.Date()`"
output: 
    pdf_document :
      latex_engine : xelatex
---

```{r setup, include=FALSE}
df1 = read.csv('BASA_AUC_2028_912.csv')
df2 = read.csv('dat_F_sub.csv')
df3 = read.csv('dat_P_sub_c.csv')
df4 = read.csv('years20262030.csv')
library(tidyr)
library(dplyr)
```

Pick dat_P_sub_c data set.

```{r cleaning, inlcude=TRUE}
sum(duplicated(df3$Pass_ID))
# check WT_flag indicator is valid
error = 0
for (i in (1:nrow(df3))){
  if (is.na(df3$Wait_Time[i]) == 'TRUE' & df3$WT_flag[i]=='0'){
    error = error + 1
    print(i)
  }else if (is.na(df3$Wait_Time[i]) == 'FALSE' & df3$WT_flag[i]=='1'){
    error = error + 1
    print(i)
  }
}
```

No duplicate passengers

```{r exploration, include=TRUE}
table(df3$Season)
table(df3$Day_of_Week)
library(ggplot2)
ggplot(df3, aes(x=Day_of_Week)) + 
  geom_bar()+
  geom_text(stat='count', aes(label=after_stat(count)), vjust=-1)
```

With Burke's theorem for M/M//c, we can estimate the arrival rate in a cluster by the departure process. We may estimate the S1's arrival rate All $S_1$ arrivals will eventually leave at $S_2$ and so the fluctuations at $S_2$ follow the same statistical property governing arrivals to the queue: arrival rates can be estimated by using data readings at $S_2$ within a given cluster.

```{r arrival rate, include=TRUE}
df = data.frame(matrix(ncol = 6, nrow = 0))
x = c("Season", "Day_of_Week", "Time_of_Day", "Total_Time", "Total_Passengers ", "Average_Servers")
colnames(df) = x

for (season in unique(df3$Season)){
  for (day in unique(df3$Day_of_Week)){
    for (time in unique(df3$Time_of_Day)){
      # total number of passengers 
      total = (df3 %>% count(Season == season & Day_of_Week == day & Time_of_Day == time, sort = TRUE))[2,2]
      # total number of hours, times 6 since each cluster contains 6 hours
      tt = length(unique((df3[df3$Season == season & df3$Day_of_Week == day & df3$Time_of_Day == time,])$Departure_Date)) * 6
      # average servers in QBS
      aver_server = round(sum(df3[which(df3$Season == season & df3$Day_of_Week == day & df3$Time_of_Day == time), 8])/total,3)
      # add to data frame
      new = c(season, day, time, tt, total, aver_server)
      df[nrow(df) + 1, ] = new 
    }
  }
}
# compute average arrival rate
df['Avg_Arrival_Rate'] = round(as.numeric(df$Total_Passengers) / as.numeric(df$Total_Time),3)
# sort columns
df = df[order(df$Season, df$Day_of_Week, df$Time_of_Day),]
# write.csv(df, "dat_p.csv", row.names=FALSE)
```
